How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?

### 2024-11-04 paper.tex
before proof, make compatible with up-to-date R version

  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}

@@

  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{csquotes}
\usepackage{placeins}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?}
\author{\phantom{0}}
\date{}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{} }

@@

In categorical choice contexts, there is ample and long-standing (e.g. Crutchfield, 1955) evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see H. Mercier \& Morin, 2019). This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role (e.g.~because the answers are private, see Asch, 1956). Similar results have been obtained with young children (Bernard, Harris, Terrier, \& Clément, 2015; Bernard, Proust, \& Clément, 2015; Chen, Corriveau, \& Harris, 2013; Corriveau, Fusaro, \& Harris, 2009; e.g. Fusaro \& Harris, 2008; Herrmann, Legare, Harris, \& Whitehouse, 2013; Morgan, Laland, \& Harris, 2015).

@@

Should we expect that people be able to infer that more convergent answers likely indicate not only more accurate answers, but also that those who gave the answers were competent? In order to make the best of communicated information, humans have to be able to evaluate it, so as to discard inaccurate or harmful information, while accepting accurate and beneficial information (Smith \& Harper, 2003). It has been argued that a suite of cognitive mechanisms--mechanisms of epistemic vigilance--evolved to serve this function (Hugo Mercier, 2020; Sperber et al., 2010). Since the opinion of more than one individual is often available to us, there should be mechanisms of epistemic vigilance dedicated to processing such situations. It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven't been formed independently of each other (Hugo Mercier \& Miton, 2019). Evidence suggests that these mechanisms rely on heuristics which become more refined with age (Morgan et al., 2015), and which are far from perfect (Hugo Mercier \& Miton, 2019; in particular, they ignore many cases of informational dependencies, see, e.g. Yousif, Aboody, \& Keil, 2019). As mentioned above, in the experiments evaluating how people process convergent information, the participants had grounds to believe that the information came from competent informants. However, the same mechanisms could lead people to perform the same inference when they do not have such information (especially since, as is shown presently, such an inference is warranted).

@@

Our models indicate that groups of informants are more likely to have given accurate answers, and to be competent, when their answers converge. In a series of experiments, we test whether people draw these inferences both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6). By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be. There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information (Desai, Xie, \& Hayes, 2022; Hugo Mercier \& Miton, 2019; Xie \& Hayes, 2022; Yin, Xu, Lin, Zhou, \& Guo, 2024; Yousif et al., 2019). To test whether participants took into account dependencies between the informants in the present context, this factor was manipulated in two different ways (in Experiment 2, some informants had discussed with each other; in Experiments 3 and 5, some informants had an incentive to provide the same answer). We predicted that dependencies between the informants' answers should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.

@@

Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently made estimates, to be less accurate, or produced by less competent individuals. This might stem from the fact that participants, in various situations, neglect informational dependencies (Yousif et al., 2019), or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks (for review, see, e.g., H. Mercier, 2016), including numerical estimates (e.g. H. Mercier \& Claidière, 2022). As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.

@@

When people see that others agree with each other, they tend to believe that they are right. This inference has been evidenced in several experiments, both for numerical estimates (Budescu \& Rantilla, 2000; Budescu et al., 2003; Budescu \& Yu, 2007; e.g. Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see H. Mercier \& Morin, 2019). However, these experiments do not test whether this inference of accuracy of the information extends to an inference of competence of the informants. Moreover, by their design, participants arguably assumed a degree of competence among the informants. For instance, when children are confronted with several individuals who agree on how to name a novel object (e.g. Corriveau et al., 2009), they can assume that these (adult) individuals tend to know what the names of objects are. If a certain competence of the informants is assumed, then well-known results from the literature on judgment aggregation--the wisdom of crowds--show that the average opinion of a set of individuals is, in a wide range of circumstances, more likely to be accurate than that of a single individual (e.g. Larrick \& Soll, 2006).

@@

The present experiments show that people draw inferences from convergence to accuracy and to competence, but then do not precisely show how they do it. As suggested in the introduction, it is plausible that participants use the combination of two heuristics: one leading them from convergence to accuracy, and one from accuracy to competence. This two step process is coherent with the observation that the inference from convergence to accuracy (one step) is stronger than that from convergence to competence (two steps). However, our results are also compatible with participants performing two inferences from convergence, one to accuracy and one to competence, the latter being weaker because the inferences roughly follow the normative model. Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants' opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account (H. Mercier \& Morin, 2019).

@@

\bibitem[\citeproctext]{ref-mercierNotBornYesterday2020}
Mercier, Hugo. (2020). \emph{Not born yesterday: the science of who we trust and what we believe}. Retrieved from \url{https://doi.org/10.1515/9780691198842}

@@

\bibitem[\citeproctext]{ref-mercierUtilizingSimpleCues2019}
Mercier, Hugo, \& Miton, H. (2019). Utilizing simple cues to informational dependency. \emph{Evolution and Human Behavior}, \emph{40}(3), 301--314. \url{https://doi.org/10.1016/j.evolhumbehav.2019.01.001}

@@

In our preregistration, we stated we would run paired t-tests and only conduct mixed models as a robustness check. However, it became clear to us later that mixed models are more appropriate in light of our data structure. In the main article, we therefore only report the results of the mixed models. Note, however, that using paired t-tests, we confirm our results (Accuracy: -1.37, t-statistic = -26.14, p \textless{} .001; Competence: -1.23, t-statistic = -21.82, p \textless{} .001)

@@

\begin{table}

\caption{\label{tab:exp1-regression-table}Model Results Experiment 1}
\centering
\begin{tabular}[t]{lcc}
\toprule
  & Accuracy & Competence\\
\midrule
(Intercept) & 3.105 & 3.516\\
 & (0.077) & (0.068)\\
convergenceconvergent & 1.368 & 1.232\\
 & (0.072) & (0.085)\\
numberlarge & 0.175 & 0.002\\
 & (0.050) & (0.048)\\
SD (Intercept ID) & 0.905 & 0.764\\
SD (convergenceconvergent ID) & 0.732 & 0.993\\
Cor (Intercept\textasciitilde{}convergenceconvergent ID) & -0.128 & -0.554\\
SD (Observations) & 1.006 & 0.955\\
\midrule
Num.Obs. & 1600 & 1600\\
R2 Marg. & 0.191 & 0.195\\
R2 Cond. & 0.594 & 0.532\\
AIC & 5131.0 & 4950.5\\
BIC & 5168.7 & 4988.1\\
ICC & 0.5 & 0.4\\
RMSE & 0.91 & 0.86\\
\bottomrule
\end{tabular}
\end{table}

@@

We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \ref{tab:exp2-conditions}).

\caption{\label{tab:exp2-conditions}}

\begin{tabular}{m{3cm}m{12cm}}
\toprule
Condition & \multicolumn{1}{c}{Description}\\
\midrule
Independence & Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.\\
Discussion & Players are asked to talk with each other about the game at length before giving their estimates.\\
\bottomrule
\end{tabular}

@@

\subsection{Design}\label{design-7}

@@

We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \ref{tab:conditions-exp3}; between participants) and convergence (two levels, convergence and divergence; within participants). Participants saw four scenarios, one for each combination of the two factors convergence and informational dependency.

\caption{\label{tab:conditions-exp3}}

\begin{tabular}{m{3cm}m{12cm}}
\toprule
Condition & \multicolumn{1}{c}{Description}\\
\midrule
Independence & Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.\\
Conflict of interest & All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future.\\
\bottomrule
\end{tabular}

@@

\begin{table}

\caption{\label{tab:exp3-table}}
\centering
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Divergent & Convergent & Divergent & Convergent\\
\midrule
Conflict of interest & 3.566 (sd = 1.215) & 4.455 (sd = 1.409) & 3.732 (sd = 1.101) & 4.556 (sd = 1.19)\\
Independent & 3.4 (sd = 1.08) & 5.28 (sd = 1.052) & 3.61 (sd = 1.111) & 5.235 (sd = 0.992)\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}

\caption{\label{tab:model-vs-participant-table}Participants vs. Model (Exp. 4)}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Participants} & \multicolumn{2}{c}{Model} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Accuracy & Competence & Accuracy  & Competence \\
\midrule
(Intercept) & 28.859*** & 3.413*** & 15.211*** & 3.233***\\
 & (1.702) & (0.098) & (0.221) & (0.008)\\
convergence & 16.838*** & 0.683*** & 28.471*** & 0.425***\\
 & (0.922) & (0.053) & (0.097) & (0.003)\\
SD (Intercept id) & 15.926 & 0.929 &  & \\
SD (convergence id) & 8.640 & 0.497 &  & \\
Cor (Intercept\textasciitilde{}convergence id) & -0.808 & -0.857 &  & \\
SD (Observations) & 10.159 & 0.553 &  & \\
\midrule
Num.Obs. & 800 & 800 & 333000 & 333000\\
R2 &  &  & 0.205 & 0.047\\
R2 Adj. &  &  & 0.205 & 0.047\\
R2 Marg. & 0.555 & 0.408 &  & \\
R2 Cond. & 0.839 & 0.786 &  & \\
AIC & 6402.6 & 1756.0 & 3301803.1 & 1051474.5\\
BIC & 6430.7 & 1784.1 & 3301835.2 & 1051506.7\\
ICC & 0.6 & 0.6 &  & \\
Log.Lik. &  &  & -1650898.534 & -525734.261\\
F &  &  & 85844.746 & 16470.902\\
RMSE & 8.98 & 0.49 & 34.42 & 1.17\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\end{tabular}
\end{table}

@@

\subsection{Design}\label{design-8}

@@

We manipulated convergence within participants in the same way we did in experiment 4. In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \ref{tab:exp5-condition}). In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is. In the independent condition, there was no such conflict of interest and experts were described as independent.

\caption{\label{tab:exp5-condition}}

\begin{tabular}{m{3cm}m{12cm}}
\toprule
Condition & \multicolumn{1}{c}{Description}\\
\midrule
Independence condition & The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.\\
Conflict of interest condition & The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations.\\
\bottomrule
\end{tabular}

@@

\begin{table}

\caption{\label{tab:exp5-table}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Conflict of interest & Independent & Conflict of interest & Independent\\
\midrule
minority (0) & 39.733 (sd = 20.238) & 36.46 (sd = 16.793) & 3.802 (sd = 1.146) & 3.606 (sd = 1.006)\\
divergence (1) & 41.52 (sd = 13.344) & 42.51 (sd = 13.6) & 4.005 (sd = 0.878) & 4.045 (sd = 0.839)\\
majority (2) & 55.416 (sd = 16.637) & 59.515 (sd = 14.575) & 4.495 (sd = 0.921) & 4.737 (sd = 0.879)\\
consensus (3) & 66.198 (sd = 24.825) & 71.914 (sd = 23.457) & 4.96 (sd = 1.273) & 5.247 (sd = 1.264)\\
\bottomrule
\end{tabular}}
\end{table}

@@

\subsection{Design}\label{design-9}

@@

\begin{table}

\caption{\label{tab:exp6-table}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & 10 options & 3 options & 10 options & 3 options\\
\midrule
minority (0) & 32.333 (sd = 22.987) & 35.703 (sd = 20.829) & 3.61 (sd = 1.245) & 3.574 (sd = 1.227)\\
divergence (1) & 37.287 (sd = 19.569) & 40.861 (sd = 18.759) & 4.067 (sd = 0.831) & 4.01 (sd = 0.881)\\
majority (2) & 63.69 (sd = 21.826) & 64.301 (sd = 17.355) & 4.957 (sd = 0.885) & 4.834 (sd = 0.881)\\
consensus (3) & 79.967 (sd = 24.15) & 80.152 (sd = 20.774) & 5.647 (sd = 1.035) & 5.466 (sd = 1.051)\\
\bottomrule
\end{tabular}}
\end{table}

@@

\begin{table}

>>>>

  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}

@@

  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\makeatother
\keywords{consensus, convergence, majority rules, competence, trust in science}
\usepackage{csquotes}
\usepackage{placeins}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?},
  pdfauthor={Jan Pfänder1, Benoît De Courson2, \& Hugo Mercier1},
  pdflang={en-EN},
  pdfkeywords={consensus, convergence, majority rules, competence, trust in science},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?}
\author{Jan Pfänder\textsuperscript{1}, Benoît De Courson\textsuperscript{2}, \& Hugo Mercier\textsuperscript{1}}
\date{}

@@

\authornote{

We thank the anonymous reviewers for their very valuable feedback. HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation (``An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference''). The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).

Correspondence concerning this article should be addressed to Hugo Mercier, . E-mail: \href{mailto:hugo.mercier@gmail.com}{\nolinkurl{hugo.mercier@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institut Jean Nicod, Département d'études cognitives, ENS, EHESS, PSL University, CNRS, France\\\textsuperscript{2} Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany}

@@

In categorical choice contexts, there is ample and long-standing (e.g. Crutchfield, 1955) evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see Mercier \& Morin, 2019). This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role (e.g.~because the answers are private, see Asch, 1956). Similar results have been obtained with young children (Bernard, Harris, Terrier, \& Clément, 2015; Bernard, Proust, \& Clément, 2015; Chen, Corriveau, \& Harris, 2013; Corriveau, Fusaro, \& Harris, 2009; e.g. Fusaro \& Harris, 2008; Herrmann, Legare, Harris, \& Whitehouse, 2013; Morgan, Laland, \& Harris, 2015).

@@

Should we expect that people be able to infer that more convergent answers likely indicate not only more accurate answers, but also that those who gave the answers were competent? In order to make the best of communicated information, humans have to be able to evaluate it, so as to discard inaccurate or harmful information, while accepting accurate and beneficial information (Smith \& Harper, 2003). It has been argued that a suite of cognitive mechanisms--mechanisms of epistemic vigilance--evolved to serve this function (Mercier, 2020; Sperber et al., 2010). Since the opinion of more than one individual is often available to us, there should be mechanisms of epistemic vigilance dedicated to processing such situations. It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven't been formed independently of each other (Mercier \& Miton, 2019). Evidence suggests that these mechanisms rely on heuristics which become more refined with age (Morgan et al., 2015), and which are far from perfect (Mercier \& Miton, 2019; in particular, they ignore many cases of informational dependencies, see, e.g. Yousif, Aboody, \& Keil, 2019). As mentioned above, in the experiments evaluating how people process convergent information, the participants had grounds to believe that the information came from competent informants. However, the same mechanisms could lead people to perform the same inference when they do not have such information (especially since, as is shown presently, such an inference is warranted).

@@

Our models indicate that groups of informants are more likely to have given accurate answers, and to be competent, when their answers converge. In a series of experiments, we test whether people draw these inferences both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6). By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be. There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information (Desai, Xie, \& Hayes, 2022; Mercier \& Miton, 2019; Xie \& Hayes, 2022; Yin, Xu, Lin, Zhou, \& Guo, 2024; Yousif et al., 2019). To test whether participants took into account dependencies between the informants in the present context, this factor was manipulated in two different ways (in Experiment 2, some informants had discussed with each other; in Experiments 3 and 5, some informants had an incentive to provide the same answer). We predicted that dependencies between the informants' answers should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.

@@

Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently made estimates, to be less accurate, or produced by less competent individuals. This might stem from the fact that participants, in various situations, neglect informational dependencies (Yousif et al., 2019), or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks (for review, see, e.g., Mercier, 2016), including numerical estimates (e.g. Mercier \& Claidière, 2022). As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.

@@

When people see that others agree with each other, they tend to believe that they are right. This inference has been evidenced in several experiments, both for numerical estimates (Budescu \& Rantilla, 2000; Budescu et al., 2003; Budescu \& Yu, 2007; e.g. Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see Mercier \& Morin, 2019). However, these experiments do not test whether this inference of accuracy of the information extends to an inference of competence of the informants. Moreover, by their design, participants arguably assumed a degree of competence among the informants. For instance, when children are confronted with several individuals who agree on how to name a novel object (e.g. Corriveau et al., 2009), they can assume that these (adult) individuals tend to know what the names of objects are. If a certain competence of the informants is assumed, then well-known results from the literature on judgment aggregation--the wisdom of crowds--show that the average opinion of a set of individuals is, in a wide range of circumstances, more likely to be accurate than that of a single individual (e.g. Larrick \& Soll, 2006).

@@

The present experiments show that people draw inferences from convergence to accuracy and to competence, but then do not precisely show how they do it. As suggested in the introduction, it is plausible that participants use the combination of two heuristics: one leading them from convergence to accuracy, and one from accuracy to competence. This two step process is coherent with the observation that the inference from convergence to accuracy (one step) is stronger than that from convergence to competence (two steps). However, our results are also compatible with participants performing two inferences from convergence, one to accuracy and one to competence, the latter being weaker because the inferences roughly follow the normative model. Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants' opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account (Mercier \& Morin, 2019).

@@

\bibitem[\citeproctext]{ref-mercierNotBornYesterday2020}
Mercier, H. (2020). \emph{Not born yesterday: the science of who we trust and what we believe}. Retrieved from \url{https://doi.org/10.1515/9780691198842}

@@

\bibitem[\citeproctext]{ref-mercierUtilizingSimpleCues2019}
Mercier, H., \& Miton, H. (2019). Utilizing simple cues to informational dependency. \emph{Evolution and Human Behavior}, \emph{40}(3), 301--314. \url{https://doi.org/10.1016/j.evolhumbehav.2019.01.001}

@@

In our preregistration, we stated we would run paired t-tests and only conduct mixed models as a robustness check. However, it became clear to us later that mixed models are more appropriate in light of our data structure. In the main article, we therefore only report the results of the mixed models. Note, however, that using paired t-tests, we confirm our results (Accuracy: 1.37, t-statistic = 26.14, p \textless{} .001; Competence: 1.23, t-statistic = 21.82, p \textless{} .001)

@@

\begin{table}
\centering
\caption{\label{tab:exp1-regression-table}Model Results Experiment 1}
\centering
\begin{tabular}[t]{lcc}
\toprule
  & Accuracy & Competence\\
\midrule
(Intercept) & 3.105 & 3.516\\
 & (0.077) & (0.068)\\
convergenceconvergent & 1.368 & 1.232\\
 & (0.072) & (0.085)\\
numberlarge & 0.175 & 0.002\\
 & (0.050) & (0.048)\\
SD (Intercept ID) & 0.905 & 0.764\\
SD (convergenceconvergent ID) & 0.732 & 0.993\\
Cor (Intercept\textasciitilde{}convergenceconvergent ID) & -0.128 & -0.554\\
SD (Observations) & 1.006 & 0.955\\
\midrule
Num.Obs. & 1600 & 1600\\
R2 Marg. & 0.191 & 0.195\\
R2 Cond. & 0.594 & 0.532\\
AIC & 5131.0 & 4950.5\\
BIC & 5168.7 & 4988.1\\
ICC & 0.5 & 0.4\\
RMSE & 0.91 & 0.86\\
\bottomrule
\end{tabular}
\end{table}

@@

@@

@@

@@

\begin{table}
\centering
\caption{\label{tab:exp3-table}}
\centering
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Divergent & Convergent & Divergent & Convergent\\
\midrule
Conflict of interest & 3.566 (sd = 1.215) & 4.455 (sd = 1.409) & 3.732 (sd = 1.101) & 4.556 (sd = 1.19)\\
Independent & 3.4 (sd = 1.08) & 5.28 (sd = 1.052) & 3.61 (sd = 1.111) & 5.235 (sd = 0.992)\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{\label{tab:model-vs-participant-table}Participants vs. Model (Exp. 4)}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Participants} & \multicolumn{2}{c}{Model} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Accuracy & Competence & Accuracy  & Competence \\
\midrule
(Intercept) & 28.859*** & 3.413*** & 15.211*** & 3.233***\\
 & (1.702) & (0.098) & (0.221) & (0.008)\\
convergence & 16.838*** & 0.683*** & 28.471*** & 0.425***\\
 & (0.922) & (0.053) & (0.097) & (0.003)\\
SD (Intercept id) & 15.926 & 0.929 &  & \\
SD (convergence id) & 8.640 & 0.497 &  & \\
Cor (Intercept\textasciitilde{}convergence id) & -0.808 & -0.857 &  & \\
SD (Observations) & 10.159 & 0.553 &  & \\
\midrule
Num.Obs. & 800 & 800 & 333000 & 333000\\
R2 &  &  & 0.205 & 0.047\\
R2 Adj. &  &  & 0.205 & 0.047\\
R2 Marg. & 0.555 & 0.408 &  & \\
R2 Cond. & 0.839 & 0.786 &  & \\
AIC & 6402.6 & 1756.0 & 3301803.1 & 1051474.5\\
BIC & 6430.7 & 1784.1 & 3301835.2 & 1051506.7\\
ICC & 0.6 & 0.6 &  & \\
Log.Lik. &  &  & -1650898.534 & -525734.261\\
RMSE & 8.98 & 0.49 & 34.42 & 1.17\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\end{tabular}
\end{table}

@@

@@

@@

\begin{table}
\centering
\caption{\label{tab:exp5-table}}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Conflict of interest & Independent & Conflict of interest & Independent\\
\midrule
minority (0) & 39.733 (sd = 20.238) & 36.46 (sd = 16.793) & 3.802 (sd = 1.146) & 3.606 (sd = 1.006)\\
divergence (1) & 41.52 (sd = 13.344) & 42.51 (sd = 13.6) & 4.005 (sd = 0.878) & 4.045 (sd = 0.839)\\
majority (2) & 55.416 (sd = 16.637) & 59.515 (sd = 14.575) & 4.495 (sd = 0.921) & 4.737 (sd = 0.879)\\
consensus (3) & 66.198 (sd = 24.825) & 71.914 (sd = 23.457) & 4.96 (sd = 1.273) & 5.247 (sd = 1.264)\\
\bottomrule
\end{tabular}}
\end{table}

@@

@@

\begin{table}
\centering\centering
\caption{\label{tab:exp6-table}}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & 10 options & 3 options & 10 options & 3 options\\
\midrule
minority (0) & 32.333 (sd = 22.987) & 35.703 (sd = 20.829) & 3.61 (sd = 1.245) & 3.574 (sd = 1.227)\\
divergence (1) & 37.287 (sd = 19.569) & 40.861 (sd = 18.759) & 4.067 (sd = 0.831) & 4.01 (sd = 0.881)\\
majority (2) & 63.69 (sd = 21.826) & 64.301 (sd = 17.355) & 4.957 (sd = 0.885) & 4.834 (sd = 0.881)\\
consensus (3) & 79.967 (sd = 24.15) & 80.152 (sd = 20.774) & 5.647 (sd = 1.035) & 5.466 (sd = 1.051)\\
\bottomrule
\end{tabular}}
\end{table}

@@

\begin{table}
\centering
\caption{\label{tab:unnamed-chunk-98}Participants vs. Model (Exp. 6)}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Participants} & \multicolumn{2}{c}{Model} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Accuracy & Competence & Accuracy  & Competence \\
\midrule
(Intercept) & 29.830*** & 3.508*** & 11.407*** & 3.189***\\
 & (1.197) & (0.063) & (0.220) & (0.009)\\
convergence & 16.305*** & 0.675*** & 31.596*** & 0.526***\\
 & (0.600) & (0.032) & (0.130) & (0.005)\\
number\_options\_effect\_code & -3.812 & 0.024 & -13.794*** & -0.207***\\
 & (2.394) & (0.126) & (0.439) & (0.018)\\
convergence × number\_options\_effect\_code & 1.252 & 0.050 & 8.498*** & 0.251***\\
 & (1.200) & (0.065) & (0.260) & (0.011)\\
SD (Intercept id) & 19.383 & 1.031 &  & \\
SD (convergence id) & 9.625 & 0.527 &  & \\
Cor (Intercept\textasciitilde{}convergence id) & -0.699 & -0.829 &  & \\
SD (Observations) & 12.103 & 0.598 &  & \\
\midrule
Num.Obs. & 2384 & 2384 & 51003 & 51003\\
R2 &  &  & 0.539 & 0.165\\
R2 Adj. &  &  & 0.539 & 0.165\\
R2 Marg. & 0.423 & 0.355 &  & \\
R2 Cond. & 0.814 & 0.779 &  & \\
AIC & 19980.3 & 5624.5 & 497742.9 & 170192.2\\
BIC & 20026.5 & 5670.7 & 497787.1 & 170236.4\\
ICC & 0.7 & 0.7 &  & \\
Log.Lik. &  &  & -248866.472 & -85091.109\\
RMSE & 10.67 & 0.53 & 31.83 & 1.28\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\end{tabular}
\end{table}
